---
title: "Part 2 Chi-square with lizards' tails"
author: "Kirsten Hodgson"
date: "12/4/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(here)
library(janitor)
library(broom)
```

## Read in data

```{r}
lizards <- read.csv(here("data","lizards.csv"))
```

Get counts of lizards by species
```{r}
lizards %>% 
  count(spp) %>%  #count observations by species
  arrange(-n) #decreasing order of count
```

## Make a subset

- UTST lizards
- Sites with the 3 highest observed populations (IBPE (enclosure grassland), RABB (rabbit mesquite site), EAST (east tarbush site))
- Tail condition is recorded (W = whole, B = broken)

```{r}
utst <- lizards %>% 
  filter(spp == "UTST") %>% 
  filter(site %in% c("IBPE", "RABB", "EAST")) %>% 
  filter(tail %in% c("B","W"))
```

## Find counts and proportions of lizards with broken or whole tails at each site

```{r}
utst %>% 
  count(site, tail) #Fine but in long format, not ideal for creating contingency table
```

```{r}
tail_counts <- utst %>% 
  janitor::tabyl(site, tail)

tail_proportions <- tail_counts %>% #all in janitor package
  adorn_percentages() %>% 
  adorn_pct_formatting(digits = 2) %>% 
  adorn_ns()
```

## Convert to a contingency table

Use `column_to_rowname()`.

```{r}
lizard_ct <- tail_counts %>% 
  column_to_rownames(var = "site")
```

## Chi-square

```{r}
lizard_x2 <- chisq.test(lizard_ct)
lizard_x2

lizard_tidy <- broom::tidy(lizard_x2)
```

There is a significant association (i.e. non-independence) between site and lizard tail breakage ($\chi$^2^(`r lizard_tidy$parameter`) = `r lizard_tidy$statistic`, *p* = `r lizard_tidy$p.value`).
